---
title: "Interactive Molasses-Methane Calculator"
format: 
  html:
    page-layout: full
    code-fold: true
server: shiny
---

```{r setup}
#| include: false
library(shiny)
library(reticulate)
```

## Introduction

This interactive calculator helps you:

1. Calculate the amount of molasses required to produce a specific amount of methane
2. Calculate the amount of CO2 and CH4 produced from a given amount of molasses

The calculations are based on the biochemical pathway:
Sucrose → Glucose → Ethanol → Methane + CO2

## Calculator

```{python}
#| context: setup
# Define molecular weights
MW_sucrose = 342.30  # g/mol
MW_glucose = 180.18  # g/mol
MW_ethanol = 46.08   # g/mol
MW_CO2 = 44.01       # g/mol
MW_CH4 = 16.04       # g/mol

# Percentage of sucrose in molasses by weight
sucrose_content = 0.50  # 50%

def calculate_molasses_for_methane(methane_kg):
    """Calculate molasses required for a specific amount of methane."""
    # Calculations
    desired_methane_g = methane_kg * 1000
    moles_CH4_required = desired_methane_g / MW_CH4
    moles_ethanol_required = (2 / 3) * moles_CH4_required
    moles_glucose_required = moles_ethanol_required / 2
    moles_sucrose_required = moles_glucose_required / 2
    mass_sucrose_required = moles_sucrose_required * MW_sucrose
    mass_molasses_required = mass_sucrose_required / sucrose_content
    mass_molasses_required_tonnes = mass_molasses_required / 1000000
    
    return mass_molasses_required_tonnes

def calculate_gases_from_molasses(molasses_tonnes):
    """Calculate the amount of CO2 and CH4 produced from molasses."""
    # Calculations
    molasses_mass_g = molasses_tonnes * 1000000
    sucrose_mass = molasses_mass_g * sucrose_content
    moles_sucrose = sucrose_mass / MW_sucrose
    moles_glucose = 2 * moles_sucrose
    moles_ethanol = 2 * moles_glucose
    moles_CH4 = 1.5 * moles_ethanol
    moles_CO2_from_glucose = 2 * moles_glucose
    moles_CO2_from_ethanol = 0.5 * moles_ethanol
    total_moles_CO2 = moles_CO2_from_glucose + moles_CO2_from_ethanol
    total_mass_CO2_kg = (total_moles_CO2 * MW_CO2) / 1000
    total_mass_CH4_kg = (moles_CH4 * MW_CH4) / 1000
    
    return total_mass_CO2_kg, total_mass_CH4_kg
```

```{r}
#| panel: sidebar
tabsetPanel(
  tabPanel("Molasses Required",
    numericInput("methane_kg", "Desired methane (kg):", value = 1.0, min = 0.01),
    actionButton("calculate1", "Calculate", class = "btn-primary"),
    hr(),
    verbatimTextOutput("molasses_result")
  ),
  tabPanel("Gas Production",
    numericInput("molasses_tonnes", "Molasses input (tonnes):", value = 1.0, min = 0.01),
    actionButton("calculate2", "Calculate", class = "btn-primary"),
    hr(),
    verbatimTextOutput("gas_result")
  )
)
```

```{r}
#| panel: fill

observeEvent(input$calculate1, {
  methane_kg <- input$methane_kg
  result <- py$calculate_molasses_for_methane(methane_kg)
  output$molasses_result <- renderText({
    paste0("Total molasses required: ", format(result, digits = 3), " metric tonnes")
  })
})

observeEvent(input$calculate2, {
  molasses_tonnes <- input$molasses_tonnes
  result <- py$calculate_gases_from_molasses(molasses_tonnes)
  output$gas_result <- renderText({
    paste0(
      "CO2 produced: ", format(result[[0]], digits = 3), " kilograms\n",
      "CH4 produced: ", format(result[[1]], digits = 3), " kilograms"
    )
  })
})
```

## Reaction Pathway

The calculator uses these simplified biochemical reactions:

1. Sucrose (C₁₂H₂₂O₁₁) → 2 Glucose (C₆H₁₂O₆)
2. Glucose (C₆H₁₂O₆) → 2 Ethanol (C₂H₅OH) + 2 CO₂
3. 2 Ethanol (C₂H₅OH) → 3 CH₄ + 1 CO₂

Note: These are simplified reactions. Actual bioprocesses may have different yields.
---
